---
title: "Medical insurance charges: Random forest model"
author: "Jeewoen Shin"
date: "02/16/2022"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
## setup
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(root.dir="...")

library(tidyverse) # tidyverse includes dplyr
library(gtsummary)
library(ranger)
library(ggplot2)
library(corrplot)
```

# Load input data and check missing data
```{r, results="hide"} 
insurance.data.full = readr::read_csv("data/insurance.csv", 
                              col_types = cols( age = col_integer(),
                              sex = col_factor(),
                              bmi = col_double(),
                              children = col_integer(),
                              smoker = col_factor(),
                              region = col_factor(),
                              charges = col_double()
));

# check missing data
sum(is.na(insurance.data.full))
# view raw data
insurance.data.full %>% tbl_summary() # reporting the median and IQR
insurance.data.full %>% tbl_summary(statistic = list(all_continuous() ~ "{mean} ({sd})")) 
# reporting mean and SD 

insurance.data <- insurance.data.full %>% mutate(id = row_number()) # person id
```

# Split into training and test set
```{r, results="hide"} 
set.seed(0)

# split training & test set
training_set = insurance.data %>% sample_frac(0.75)
test_set = anti_join(insurance.data, training_set, by='id')

training_set = training_set %>% select(-"id")
test_set = test_set %>% select(-"id")
```

# Build random forest model (ranger)
```{r, results="hide"} 
set.seed(1)
# importance = permutation: 
rf.model.permutation <- ranger(charges ~ ., data = training_set, 
                               importance = "permutation", mtry=3)

if(FALSE){ #importance = impurity: variance of the responses for regression 
  rf.model.impurity <- ranger(cases ~ ., data = training_set, 
                              importance = "impurity", mtry=3)
}
``` 
Note: mtry = the number of variables randomly sampled as candidates at each split. Default: floor(# variables/3) which is 2. Increase mtry to 3 instead of 2.

## Print R squared and MSE
```{r, tidy = TRUE} 
print(rf.model.permutation)
#print(rf.model.impurity)
```
## Print variable importance
```{r, tidy = TRUE}
rf.model.permutation$variable.importance
#rf.model.impurity$variable.importance
```
## Plot variable importance
```{r, tidy = TRUE} 
varImportance.permutation = data.frame(Variables = names(rf.model.permutation$variable.importance), 
                                    Importance =round(rf.model.permutation$variable.importance,2))

rankImportance.permutation=varImportance.permutation%>%mutate(Rank=paste('#',dense_rank(desc(Importance))))

ggplot(rankImportance.permutation,aes(x=reorder(Variables,Importance),
                                   y=Importance,fill=Importance))+ 
  geom_bar(stat='identity') + 
  geom_text(aes(x = Variables, y = 0.5, label = Rank),
            hjust=0, vjust=0.55, size = 4, colour = 'red') +
  labs(x = 'Variables') +
  coord_flip() + 
  theme_classic()
```
```{r, include=FALSE}
if(FALSE){
  varImportance.impurity = data.frame(Variables = names(rf.model.impurity$variable.importance), 
                             Importance =round(rf.model.impurity$variable.importance,2))
  
  rankImportance.impurity=varImportance.impurity%>%mutate(Rank=paste('#',dense_rank(desc(Importance))))
  
  ggplot(rankImportance.impurity,aes(x=reorder(Variables,Importance),
                            y=Importance,fill=Importance))+ 
    geom_bar(stat='identity') + 
    geom_text(aes(x = Variables, y = 0.5, label = Rank),
              hjust=0, vjust=0.55, size = 4, colour = 'red') +
    labs(x = 'Variables') +
    coord_flip() + 
    theme_classic()
}
```
## Model accuracy, Goodness-of-fit 
```{r, tidy = TRUE}
training_set$predicted.rf.permutation <- predict(rf.model.permutation, dat = training_set)$predictions

test_set$predicted.rf.permutation <- predict(rf.model.permutation, dat = test_set)$predictions

# calculate root-mean-square deviation
rmse <- function(actual, estimate) {
  rmse = sqrt(sum((actual - estimate)^2) / length(actual))
  return(rmse)
}

print(rmse(training_set$charges, training_set$predicted.rf.permutation)) # training set
print(rmse(test_set$charges, test_set$predicted.rf.permutation)) # test set 

if(FALSE){ # impurity
  training_set$predicted.rf.impurity <- predict(rf.model.impurity, dat = training_set)$predictions
  sum((training_set$charges - training_set$predicted.rf.impurity)^2) / nrow(training_set)
  
  test_set$predicted.rf.impurity <- predict(rf.model.impurity, dat = test_set)$predictions
  sum((test_set$charges - test_set$predicted.rf.impurity)^2) / nrow(test_set)
}
```

# Plot charges and predicted charges
```{r, tidy = TRUE}
par(mfrow=c(1,2))
plot(training_set$charges ~ training_set$predicted.rf.permutation, asp=1, pch=20, xlab="Fitted charges", ylab="Actual charges", xlim=c(0,65000), ylim=c(0,65000), main="Training set")
grid()
abline(0,1)

plot(test_set$charges ~ test_set$predicted.rf.permutation, asp=1, pch=20, xlab="Fitted charges", ylab="Actual charges", xlim=c(0,65000), ylim=c(0,65000), main="Test set")
grid()
abline(0,1)

```